When does vapor pressure deficit drive or reduce evapotranspiration?

### 2018-01-19 doc/paper/vpd_et_paper_PG_v1.tex
done with morn writing

ALL OF THE REST SHOULDB EB EXPLAINED IN MUCH MORE DETAIL DESCRIBE EACH FIGURE AND THE RESULTS For DBF and CSH our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included.  The plots for DBF and CSH look as one would expect if one just added noise on top of our theory.

The theory does not match as well for CRO and GRA. Perhaps this should be expected. Our theory, for practical reasons, does not include important sources of variability (particularly between sites) for CRO and GRA. We do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis, or differences in irrigation practices between the sites. Given that the our theory does not do as well for CRO and GRA, we will refrain from firm statements about how CRO and GRA respond to changes in VPD.

@@

While the above discussion shows that our theory has some limitations when applied to some sites, especially for crops and grasslands, another question could be, how does our theory's performance compare to often-used PET? Again, as in Section \ref{sign_func}, the PET response to changes in VPD will be constant at about 3.5. Even with uncertainty included, no values of the response approach the PET case YOU NEED TO ELABORATE WAY MORE AND GUDE THE READER THROUGH YOUR THOUGHTS. PET's status outside the error bars of our theory highlight the usefulness of our theory (and Equation \ref{et}) for estimating land response and moisture budgeting.

@@

Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; VPD}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary, including uncertainty. Each column represents a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, and each row is a different PFT. Vertical black lines denote where our theory says $VPD_{crit}$ should be. For any observations to the left of this line $\frac{\partial \; ET}{\partial \; VPD}$ should be negative (plant response dominates), and any observations the right of the line should be positive (atmospheric demand dominates). In addition to seeing how our theoretically-derived VPD threshold, $VPD_{crit}$, matches observations, these analysis DO NOT SAY PLOT OR IT SOUNDS LIKE A MSC PROJECT are also useful for observing how the scaling term affects the magnitude of the derivative, if at all.

The scatter plots generally confirm expectations from Figure \ref{agu_fig}. By comparing different columns of Figure \ref{real} we can evaluate the impact of $\sigma$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; VPD}$. $g_a$'s scaling (included in columns 1 and 3) alters the magnitude considerably, particularly between PFTs. $\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; VPD}$, as observed in Figure \ref{agu_fig}. However, the general story with the noise appears to match the cleaner signal when $\sigma$ is help constant and $VPD_{crit}$ is clearly visible. The exceptions are, as discussed above, for GRA and CRO, where the full complexity plots (Columns 1 and 2) do not match well with $\sigma$ held fixed (Columns 3 and 4).

@@

HERE DISCUSS LIMITATION OF THE METHODS TO BVERY DRY CONDITIONS AND THE FEEDBCAK TO THE ATMOSPHERE OF VERY DRY SOIL CONDITIONS (BERG ET AL 2016, GENTINE ET AL> 2016)

>>>>

For DBF our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included.  The plots for DBF look as one would expect if one just added noise on top of our theory.

For CSH our theory is successful towards this goal: the theory matches the leading order behavior of the function when uncertainty is included.  The plots for CSH look as one would expect if one just added noise on top of our theory.

The theory does not match as well for CRO . Perhaps this should be expected. Our theory, for practical reasons, does not include important sources of variability (particularly between sites) for CRO. We do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis, or differences in irrigation practices between the sites. Given that the our theory does not do as well for CRO, we will refrain from firm statements about how CRO respond to changes in VPD.

The theory does not match as well for GRA. Perhaps this should be expected. Our theory, for practical reasons, does not include important sources of variability (particularly between sites) for GRA. We do not account for seasonal changes in LAI, variability in crop height and surface roughness, differences in C3 vs C4 photosynthesis, or differences in irrigation practices between the sites. Given that the our theory does not do as well for GRA, we will refrain from firm statements about how  GRA respond to changes in VPD.

@@

While the above discussion shows that our theory has some limitations when applied to some sites, especially for crops and grasslands, another question could be, how does our theory's performance compare to often-used PET? Again, as in Section \ref{sign_func}, the PET response to changes in VPD will be constant at about 3.5. In Figure \ref{agu_fig} the ET reponse to shifts in dryness is always far below (more negative) 3.5 for observations, including uncertainty. So the PET reponse to atmospheric dryness is outside the range of all observable responses. This casts doubt on PET's usefulness for capturing the leading order behavior of water budgeting at the land surface. However our theory (Equation \ref{et}) could be used with published $uWUE$ and $g1$ as an alternative to PET to estimating land response and moisture budgeting.

@@

Figure \ref{real} presents calculated $\frac{\partial \; ET}{\partial \; VPD}$ where, unless otherwise noted, all variables in Equation \ref{d_et} are allowed to vary, including uncertainty. Each column represents a different quantity related to $\frac{\partial \; ET}{\partial \; VPD}$, and each row is a different PFT. Vertical black lines denote where our theory says $VPD_{crit}$ should be. For any observations to the left of this line $\frac{\partial \; ET}{\partial \; VPD}$ should be negative (plant response dominates), and any observations the right of the line should be positive (atmospheric demand dominates). In addition to seeing how our theoretically-derived VPD threshold, $VPD_{crit}$, matches observations, these analysis are also useful for observing how the scaling term affects the magnitude of the derivative, if at all.

The scatter figures generally confirm expectations from Figure \ref{agu_fig}. By comparing different columns of Figure \ref{real} we can evaluate the impact of $\sigma$ and $g_a$ on the variability of $\frac{\partial \; ET}{\partial \; VPD}$. $g_a$'s scaling (included in columns 1 and 3) alters the magnitude considerably, particularly between PFTs. $\sigma$ variability (included in columns 1 and 2) adds a lot of additional noise to the signal of $\frac{\partial \; ET}{\partial \; VPD}$, as observed in Figure \ref{agu_fig}. However, the general story with the noise appears to match the cleaner signal when $\sigma$ is help constant and $VPD_{crit}$ is clearly visible. The exceptions are, as discussed above, for GRA and CRO, where the full complexity plots (Columns 1 and 2) do not match well with $\sigma$ held fixed (Columns 3 and 4).

@@

\section{Limitations of theory}
\explain{HERE DISCUSS LIMITATION OF THE METHODS TO BVERY DRY CONDITIONS AND THE FEEDBCAK TO THE ATMOSPHERE OF VERY DRY SOIL CONDITIONS (BERG ET AL 2016, GENTINE ET AL> 2016)}
